### 2021-09-29 docs/paper/sections/6_implementation.tex
Complete implementation

\subsection{Data collection}
\label{sec:implementation/data_collection}

Foo!

\subsection{Usage mining}
\label{sec:implementation/usage_mining}

Foo!

@@

To support developers in answering the questions raised in \cref{sec:framework}, we have designed the following views for our extension:

\begin{enumerate}[label=(\roman*)]
	\item The \emph{dependency browser} allows to explore all downstream dependencies and, grouped for each dependency, all their references to the target package.
	\item The \emph{usage browser} displays all public package member and, grouped for each member, all dependencies and their references to this member.
	\item The \emph{CodeLens integration} provides quick access to a slice of the usage browser that is attached to the definition of each package member right in the source code editor.
\end{enumerate}

Both references and members are organized in a tree view reflecting the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm felling lucky'' button for every view that redirects the user to a random dependency resp. reference to gain a faster, unbiased impression of usage samples.

Foo!

>>>>

In our implementation, we are restraining ourselves to analyzing node.js projects written in JavaScript.
This decision is based on the large prevalence of JavaScript\footnote{GitHut 2.0: A small place to discover languages in GitHub. \url{https://madnight.github.io/githut/\#/pushes}} as well as the first-rate extent of the npm ecosystem\footnote{Modulecounts: \url{http://www.modulecounts.com/}} which has been unsurpassed since 2015.
In the following, we are describing how we apply the single steps of the approach presented above to this particular domain.

\subsection{Downstream dependency collection}
\label{sec:implementation/dependency_collection}

The de facto standard package repository for node.js packages is npm.\footnote{\url{https://www.npmjs.com/}}
To fetch downstream dependency packages from the npm registry, we utilize the npm package npm-dependants\footnote{\url{https://www.npmjs.com/package/npm-dependants}} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, in particular including a URL to the source code archive of the latest package version (commonly referred to as \emph{tarball}), by using the package-metadata package.\footnote{\url{https://www.npmjs.com/package/package-metadata}}
Finally, we download and extract each source code archive from the npm registry with the help of the download-package-tarball package.\footnote{\url{https://www.npmjs.com/package/download-package-tarball}}

\begin{lstlisting}[language=json,
	label=sec:implementation/package.json,
	caption={An example \code{package.json} manifest file for the \code{ripple-lib} package specifying multiple dependencies (truncated).},
	float=tpb,
	floatplacement=tbp,
	abovecaptionskip=-5pt
]
{
	"name": "ripple-lib",
	"version": "1.10.0",
	"description": "A TypeScript/JavaScript API for interacting with the XRP Ledger",
	"dependencies": {
		"bignumber.js": "^9.0.0",
		"https-proxy-agent": "^5.0.0",
		"jsonschema": "1.2.2"
	}
}
\end{lstlisting}

For the second approach, using a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a package.json metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a package \code{<pkg>}, we can search all available package.json files for the substring \code{"<pkg>":}.
Refined with some additional constraints to filter out package copies from \code{node\_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved, our final Sourcegraph query has the following form:

{
	\lstset{frame=none,
		aboveskip=3mm,
		belowskip=3mm,
		xleftmargin=3mm,
		xrightmargin=3mm
	}
	\begin{lstlisting}[language=sourcegraph]
"<pkg>":
	file:package\.json
	-file:node_modules/
	select:file
	count:<limit>
	\end{lstlisting}
}

We submit this search query by sending it to the publicly available Sourcegraph GraphQL API.\footnote{\url{https://docs.sourcegraph.com/api/graphql}}
For each hit, we request additional information from the API about the matched repository, including the name of the depending package and the URL to the original repository.
We use the npm package graphql-request\footnote{\url{https://www.npmjs.com/package/graphql-request}} for accessing the GraphQL API.
After retrieving a search result, we download a snapshot of the full repository by using the download-git-repo package\footnote{\url{https://www.npmjs.com/package/download-git-repo}} which combines access to the APIs of different code host providers such as GitHub or BitBucket.

Additionally, we enrich the metadata of each package with a few metrics indicating its popularity from the associated GitHub repository if specified.
The GitHub API is accessed via the official octokit library.\footnote{\url{https://octokit.github.io/}}

\subsection{Usage sample mining}
\label{sec:implementation/usage_mining}

In our solution, we are focusing on mining usage samples from downstream dependencies that are written in JavaScript or TypeScript.
\TODO{Is this the right place to introduce TypeScript?}
TypeScript is an extension of JavaScript that adds strong typing to the former which is based on a combination of implicit type inference and explicit type declarations.
As the type-checking capabilities of the TypeScript compiler are optional and can be turned off via configuration, we consider the JavaScript syntax a subset of TypeScript syntax.

For parsing the dependencies, we utilize the public TypeScript Compiler API.\footnote{\url{https://github.com/Microsoft/TypeScript/wiki/Using-the-Compiler-API}}
The TypeScript compiler's \emph{binder} component annotates every node of the parsed program with a type symbol if it can resolve one.
After performing a type analysis of the dependency by instrumenting the binder, we collect all nodes from the type-annotated ASF that refer to the target package by traversing it with a visitor function.
This visitor function matches each node against a list of defined patterns (similar to \cref{fig:usage_mining/patterns}) based on the \code{syntaxKind}s of the node and its ancestors and descendants and then looks up the declaration of the node's type symbol to compare it with the source directory of the target package.

Normally, foreign types can only be resolved if they are declared in a dependency module that has been installed into the \code{node\_modules} folder of the repository.
Because downloading and installing all dependency modules for every downstream repository of the target package would increase the computational resources of the usage mining drastically and eventually violate \cref{req1}, we skip this step but instead modify the module resolution strategy of the TypeScript compiler host.
To do so, we extend the \code{paths} parameter from the default compiler options\footnote{\url{https://www.typescriptlang.org/tsconfig\#paths}} and insert a new entry that remaps accesses to the target package name to the separate source code directory of the target package.

@@

\begin{figure*}[b]
	\centering
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\textwidth]{sections/6_implementation/extension/dependencies.png}
		\caption[LoF entry]{The dependency browser.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\textwidth]{sections/6_implementation/extension/references.png}
		\caption[LoF entry]{The usage browser.}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\textwidth]{sections/6_implementation/extension/codelens.png}
		\caption[LoF entry]{The CodeLens integration for the class \code{GraphQLError}.}
	\end{subfigure}

	\caption{Screenshots of our Visual Studio Code extension used to explore downstream dependencies of the npm package graphql.}
	\label{fig:implementation/presentation/screenshots}
\end{figure*}

To support developers in answering the questions raised in \cref{sec:framework}, we have designed three key views for our extension (see \cref{fig:implementation/presentation/screenshots}):

\begin{enumerate}[label=(\roman*)]
	\item The \emph{dependency browser} allows to explore all downstream dependencies and, grouped for each dependency, all their references to the target package.
	\item The \emph{usage browser} displays all public package members and, grouped for each member, all dependencies and their references to this member.
	\item The \emph{CodeLens integration} provides quick access to a slice of the usage browser that is attached to the definition of each package member right in the source code editor.
\end{enumerate}

Both references and members are organized in a tree view reflecting the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm feeling lucky'' button for every view that redirects the user to a random dependency resp. reference to gain a faster, unbiased impression of usage samples.

We implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the Tree View API and the CodeLens API.
One challenge was to deliver fast results and not to block the user interface because even though the lightweight mining approaches we chose to satisfy \cref{req1} above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation/TODO}).
To overcome this issue, we use pagination wherever possible and applied the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our data model using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting inside the data collection process.
